### 2016-11-14 documents/DefiniteExtensions/dep.tex
Major sync issues with Github, tried to pull from repository without committing. May cause errors. Sorry!

% === Revtex Declaration ===
\documentclass[aps, 10pt, english, twoside, pra, nofootinbib, longbibliography]{revtex4-1}

% === All of the Packages I use frequently ===
\usepackage{../../packages/document_config}
\usepackage{../../packages/shared}
\usepackage{../../packages/misc_commands}

\begin{document}
    \title{The Definite Extension Procedure for Large-Scale Marginal Satisfiability}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada \\ University of Waterloo, Waterloo, Ontario, Canada}
    % \author{Elie Wolfe}
    % \email{ewolfe@perimeter@institute.ca}
    % \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada}
    \date{\today}
    \begin{abstract}
        \todo[TC]{Write an abstract}
    \end{abstract}
    \maketitle
    \tableofcontents

    \section{Introduction}
        \begin{itemize}
            \item \todo[TC]{Marginal problem}
            \item \todo[TC]{Use cases for the marginal problem}
            \item \todo[TC]{Motivate issue of computational complexity}
            \item \todo[TC]{Will be presenting technique capable of fully solving marginal problem}
            \item \todo[TC]{Early termination offers partial solution}
        \end{itemize}

    \section{Marginal Problem}
        \begin{itemize}
            \item \todo[TC]{Set up notation and definitions}
        \end{itemize}

    \section{Existing Methods}
        \begin{itemize}
            \item \todo[TC]{Geometric interpretation of marginal Polytope}
            \item \todo[TC]{Fourier Motzkin, Convex Hull}
            \item \todo[TC]{ESP, Symmetric, etc...}
            \item \todo[TC]{Existing methods do not exploit features of interest}
            \item \todo[TC]{Logical Contextuality \& Hardy Transversals}
        \end{itemize}

    \section{Definite Extension Procedure}
        \begin{itemize}
            \item \todo[TC]{Prove features of complete solution}
            \item \todo[TC]{Homogeneous, linear, integral}
            \item \todo[TC]{Multi-set representation of inequality (certificates)}
        \end{itemize}
    \subsection{Hypergraph Transversals}
        \begin{itemize}
            \item \todo[TC]{Generalizing Logical Contextuality inequalities to more exotic antecedents}
            \item \todo[TC]{Algorithms}
        \end{itemize}

    \subsection{Irreducibility \& Facets}
        \begin{itemize}
            \item \todo[TC]{Define facets and their properties}
            \item \todo[TC]{Algorithms}
            \item \todo[TC]{Visualization}
        \end{itemize}

    \subsection{Accelerations \& Symmetries}
        \todo[TC]{**Need to prove/disprove complete symmetry group**}
        \begin{itemize}
            \item \todo[TC]{Complete Characterization of Marginal Symmetries}
            \item \todo[TC]{Outcome Symmetries}
            \item \todo[TC]{Variable Symmetries}
            \item \todo[TC]{Search space and Output massively reduced}
        \end{itemize}

    \subsection{Hierarchy vs. Relaxation}
        \todo[TC]{**Need to determine upper limit**}
        \begin{itemize}
            \item \todo[TC]{Implementation as Hierarchy or Relaxation}
        \end{itemize}

    \subsection{Convex Hull Problem}
        \begin{itemize}
            \item \todo[TC]{Possible to use this on generic convex hull}
        \end{itemize}

    \subsection{Targeted Definite Extension Procedure}
        \begin{itemize}
            \item \todo[TC]{With known incompatible distributions can target output}
            \item \todo[TC]{Seeds inequalities}
        \end{itemize}

    \section{Demonstrations}

    \subsection{Generic Problem}
    \subsection{Quantum Non-locality}
    \subsection{Something Else}

    \section{Conclusions}
        \begin{itemize}
            \item \todo[TC]{In situations where Fourier Motzkin is intractable/will never finish, other methods need to be used}
            \item \todo[TC]{If number of facets is massive, partial solution with flexibility is desired}
            \item \todo[TC]{Symmetries drastically reduce the output space and search space}
            \item \todo[TC]{DEP offers a canonical presentation of output space}
        \end{itemize}

    \section*{Acknowledgments}
    \begin{itemize}
        \item \todo[TC]{Elie}
        \item \todo[TC]{Perimeter}
        \item \todo[TC]{University of Waterloo}
        \item \todo[TC]{Possible due to Mike Lazaridis Scholarship}
    \end{itemize}

\end{document}