### 2016-11-14 figures/inequalities/nosymwitness_ineq_print.tex

\begin{align*}
	P_{ABC}(110)P_{ABC}(223) + P_{ABC}(110)P_{ABC}(233) + P_{ABC}(110)P_{ABC}(323) + P_{ABC}(110)P_{ABC}(333)
\end{align*}
\[\leq\]
\begin{align*}
	2P_{ABC}(020)P_{ABC}(213) + 2P_{ABC}(023)P_{ABC}(210) + 2P_{ABC}(023)P_{ABC}(310) + 2P_{ABC}(030)P_{ABC}(213) + 2P_{ABC}(033)P_{ABC}(210)+ \\
	2P_{ABC}(033)P_{ABC}(310) + 2P_{ABC}(120)P_{ABC}(213) + 2P_{ABC}(123)P_{ABC}(210) + 2P_{ABC}(123)P_{ABC}(310) + 2P_{ABC}(130)P_{ABC}(213)+ \\
	2P_{ABC}(132)P_{ABC}(311) + 2P_{ABC}(133)P_{ABC}(210) + 2P_{ABC}(133)P_{ABC}(310) + P_{ABC}(000)P_{ABC}(003) + P_{ABC}(000)P_{ABC}(013)+ \\
	P_{ABC}(000)P_{ABC}(023) + P_{ABC}(000)P_{ABC}(033) + P_{ABC}(000)P_{ABC}(103) + P_{ABC}(000)P_{ABC}(113) + P_{ABC}(000)P_{ABC}(123)+ \\
	P_{ABC}(000)P_{ABC}(133) + P_{ABC}(000)P_{ABC}(203) + P_{ABC}(000)P_{ABC}(213) + P_{ABC}(000)P_{ABC}(223) + P_{ABC}(003)P_{ABC}(010)+ \\
	P_{ABC}(003)P_{ABC}(020) + P_{ABC}(003)P_{ABC}(030) + P_{ABC}(003)P_{ABC}(100) + P_{ABC}(003)P_{ABC}(110) + P_{ABC}(003)P_{ABC}(120)+ \\
	P_{ABC}(003)P_{ABC}(130) + P_{ABC}(003)P_{ABC}(200) + P_{ABC}(003)P_{ABC}(210) + P_{ABC}(003)P_{ABC}(220) + P_{ABC}(003)P_{ABC}(230)+ \\
	P_{ABC}(003)P_{ABC}(300) + P_{ABC}(003)P_{ABC}(310) + P_{ABC}(003)P_{ABC}(320) + P_{ABC}(003)P_{ABC}(330) + P_{ABC}(010)P_{ABC}(013)+ \\
	P_{ABC}(010)P_{ABC}(023) + P_{ABC}(010)P_{ABC}(033) + P_{ABC}(010)P_{ABC}(103) + P_{ABC}(010)P_{ABC}(113) + P_{ABC}(010)P_{ABC}(123)+ \\
	P_{ABC}(010)P_{ABC}(133) + P_{ABC}(010)P_{ABC}(203) + P_{ABC}(010)P_{ABC}(213) + P_{ABC}(010)P_{ABC}(223) + P_{ABC}(013)P_{ABC}(020)+ \\
	P_{ABC}(013)P_{ABC}(030) + P_{ABC}(013)P_{ABC}(100) + P_{ABC}(013)P_{ABC}(110) + P_{ABC}(013)P_{ABC}(120) + P_{ABC}(013)P_{ABC}(130)+ \\
	P_{ABC}(013)P_{ABC}(200) + P_{ABC}(013)P_{ABC}(210) + P_{ABC}(013)P_{ABC}(220) + P_{ABC}(013)P_{ABC}(230) + P_{ABC}(013)P_{ABC}(300)+ \\
	P_{ABC}(013)P_{ABC}(310) + P_{ABC}(013)P_{ABC}(320) + P_{ABC}(013)P_{ABC}(330) + P_{ABC}(020)P_{ABC}(023) + P_{ABC}(020)P_{ABC}(033)+ \\
	P_{ABC}(020)P_{ABC}(103) + P_{ABC}(020)P_{ABC}(113) + P_{ABC}(020)P_{ABC}(123) + P_{ABC}(020)P_{ABC}(133) + P_{ABC}(020)P_{ABC}(203)+ \\
	P_{ABC}(020)P_{ABC}(210) + P_{ABC}(020)P_{ABC}(211) + P_{ABC}(020)P_{ABC}(212) + P_{ABC}(020)P_{ABC}(223) + P_{ABC}(020)P_{ABC}(233)+ \\
	P_{ABC}(020)P_{ABC}(310) + P_{ABC}(020)P_{ABC}(311) + P_{ABC}(020)P_{ABC}(312) + P_{ABC}(020)P_{ABC}(313) + P_{ABC}(021)P_{ABC}(210)+ \\
	P_{ABC}(021)P_{ABC}(211) + P_{ABC}(021)P_{ABC}(212) + P_{ABC}(021)P_{ABC}(213) + P_{ABC}(021)P_{ABC}(310) + P_{ABC}(021)P_{ABC}(311)+ \\
	P_{ABC}(021)P_{ABC}(312) + P_{ABC}(021)P_{ABC}(313) + P_{ABC}(022)P_{ABC}(210) + P_{ABC}(022)P_{ABC}(212) + P_{ABC}(022)P_{ABC}(213)+ \\
	P_{ABC}(022)P_{ABC}(310) + P_{ABC}(022)P_{ABC}(311) + P_{ABC}(022)P_{ABC}(312) + P_{ABC}(022)P_{ABC}(313) + P_{ABC}(023)P_{ABC}(030)+ \\
	P_{ABC}(023)P_{ABC}(100) + P_{ABC}(023)P_{ABC}(110) + P_{ABC}(023)P_{ABC}(120) + P_{ABC}(023)P_{ABC}(130) + P_{ABC}(023)P_{ABC}(200)+ \\
	P_{ABC}(023)P_{ABC}(211) + P_{ABC}(023)P_{ABC}(212) + P_{ABC}(023)P_{ABC}(213) + P_{ABC}(023)P_{ABC}(220) + P_{ABC}(023)P_{ABC}(230)+ \\
	P_{ABC}(023)P_{ABC}(300) + P_{ABC}(023)P_{ABC}(311) + P_{ABC}(023)P_{ABC}(312) + P_{ABC}(023)P_{ABC}(313) + P_{ABC}(023)P_{ABC}(320)+ \\
	P_{ABC}(023)P_{ABC}(330) + P_{ABC}(030)P_{ABC}(033) + P_{ABC}(030)P_{ABC}(103) + P_{ABC}(030)P_{ABC}(113) + P_{ABC}(030)P_{ABC}(123)+ \\
	P_{ABC}(030)P_{ABC}(133) + P_{ABC}(030)P_{ABC}(203) + P_{ABC}(030)P_{ABC}(210) + P_{ABC}(030)P_{ABC}(211) + P_{ABC}(030)P_{ABC}(212)+ \\
	P_{ABC}(030)P_{ABC}(223) + P_{ABC}(030)P_{ABC}(233) + P_{ABC}(030)P_{ABC}(310) + P_{ABC}(030)P_{ABC}(311) + P_{ABC}(030)P_{ABC}(312)+ \\
	P_{ABC}(030)P_{ABC}(313) + P_{ABC}(031)P_{ABC}(210) + P_{ABC}(031)P_{ABC}(211) + P_{ABC}(031)P_{ABC}(212) + P_{ABC}(031)P_{ABC}(213)+ \\
	P_{ABC}(031)P_{ABC}(310) + P_{ABC}(031)P_{ABC}(311) + P_{ABC}(031)P_{ABC}(312) + P_{ABC}(031)P_{ABC}(313) + P_{ABC}(032)P_{ABC}(210)+ \\
	P_{ABC}(032)P_{ABC}(212) + P_{ABC}(032)P_{ABC}(213) + P_{ABC}(032)P_{ABC}(310) + P_{ABC}(032)P_{ABC}(311) + P_{ABC}(032)P_{ABC}(312)+ \\
	P_{ABC}(032)P_{ABC}(313) + P_{ABC}(033)P_{ABC}(100) + P_{ABC}(033)P_{ABC}(110) + P_{ABC}(033)P_{ABC}(120) + P_{ABC}(033)P_{ABC}(130)+ \\
	P_{ABC}(033)P_{ABC}(200) + P_{ABC}(033)P_{ABC}(211) + P_{ABC}(033)P_{ABC}(212) + P_{ABC}(033)P_{ABC}(213) + P_{ABC}(033)P_{ABC}(220)+ \\
	P_{ABC}(033)P_{ABC}(230) + P_{ABC}(033)P_{ABC}(300) + P_{ABC}(033)P_{ABC}(311) + P_{ABC}(033)P_{ABC}(312) + P_{ABC}(033)P_{ABC}(313)+ \\
	P_{ABC}(033)P_{ABC}(320) + P_{ABC}(033)P_{ABC}(330) + P_{ABC}(100)P_{ABC}(103) + P_{ABC}(100)P_{ABC}(113) + P_{ABC}(100)P_{ABC}(123)+ \\
	P_{ABC}(100)P_{ABC}(133) + P_{ABC}(100)P_{ABC}(203) + P_{ABC}(100)P_{ABC}(213) + P_{ABC}(100)P_{ABC}(223) + P_{ABC}(103)P_{ABC}(110)+ \\
	P_{ABC}(103)P_{ABC}(120) + P_{ABC}(103)P_{ABC}(130) + P_{ABC}(103)P_{ABC}(200) + P_{ABC}(103)P_{ABC}(210) + P_{ABC}(103)P_{ABC}(220)+ \\
	P_{ABC}(103)P_{ABC}(230) + P_{ABC}(103)P_{ABC}(300) + P_{ABC}(103)P_{ABC}(310) + P_{ABC}(103)P_{ABC}(320) + P_{ABC}(103)P_{ABC}(330)+ \\
	P_{ABC}(110)P_{ABC}(113) + P_{ABC}(110)P_{ABC}(123) + P_{ABC}(110)P_{ABC}(133) + P_{ABC}(110)P_{ABC}(203) + P_{ABC}(110)P_{ABC}(213)+ \\
	P_{ABC}(110)P_{ABC}(223) + P_{ABC}(113)P_{ABC}(120) + P_{ABC}(113)P_{ABC}(130) + P_{ABC}(113)P_{ABC}(200) + P_{ABC}(113)P_{ABC}(210)+ \\
	P_{ABC}(113)P_{ABC}(220) + P_{ABC}(113)P_{ABC}(230) + P_{ABC}(113)P_{ABC}(300) + P_{ABC}(113)P_{ABC}(310) + P_{ABC}(113)P_{ABC}(320)+ \\
	P_{ABC}(113)P_{ABC}(330) + P_{ABC}(120)P_{ABC}(123) + P_{ABC}(120)P_{ABC}(133) + P_{ABC}(120)P_{ABC}(203) + P_{ABC}(120)P_{ABC}(210)+ \\
	P_{ABC}(120)P_{ABC}(211) + P_{ABC}(120)P_{ABC}(212) + P_{ABC}(120)P_{ABC}(223) + P_{ABC}(120)P_{ABC}(233) + P_{ABC}(120)P_{ABC}(310)+ \\
	P_{ABC}(120)P_{ABC}(311) + P_{ABC}(120)P_{ABC}(312) + P_{ABC}(120)P_{ABC}(313) + P_{ABC}(121)P_{ABC}(210) + P_{ABC}(121)P_{ABC}(211)+ \\
	P_{ABC}(121)P_{ABC}(212) + P_{ABC}(121)P_{ABC}(213) + P_{ABC}(121)P_{ABC}(310) + P_{ABC}(121)P_{ABC}(311) + P_{ABC}(121)P_{ABC}(312)+ \\
	P_{ABC}(121)P_{ABC}(313) + P_{ABC}(122)P_{ABC}(210) + P_{ABC}(122)P_{ABC}(212) + P_{ABC}(122)P_{ABC}(213) + P_{ABC}(122)P_{ABC}(310)+ \\
	P_{ABC}(122)P_{ABC}(311) + P_{ABC}(122)P_{ABC}(312) + P_{ABC}(122)P_{ABC}(313) + P_{ABC}(123)P_{ABC}(130) + P_{ABC}(123)P_{ABC}(200)+ \\
	P_{ABC}(123)P_{ABC}(211) + P_{ABC}(123)P_{ABC}(212) + P_{ABC}(123)P_{ABC}(213) + P_{ABC}(123)P_{ABC}(220) + P_{ABC}(123)P_{ABC}(230)+ \\
	P_{ABC}(123)P_{ABC}(300) + P_{ABC}(123)P_{ABC}(311) + P_{ABC}(123)P_{ABC}(312) + P_{ABC}(123)P_{ABC}(313) + P_{ABC}(123)P_{ABC}(320)+ \\
	P_{ABC}(123)P_{ABC}(330) + P_{ABC}(130)P_{ABC}(133) + P_{ABC}(130)P_{ABC}(203) + P_{ABC}(130)P_{ABC}(210) + P_{ABC}(130)P_{ABC}(211)+ \\
	P_{ABC}(130)P_{ABC}(212) + P_{ABC}(130)P_{ABC}(223) + P_{ABC}(130)P_{ABC}(233) + P_{ABC}(130)P_{ABC}(310) + P_{ABC}(130)P_{ABC}(311)+ \\
	P_{ABC}(130)P_{ABC}(312) + P_{ABC}(130)P_{ABC}(313) + P_{ABC}(131)P_{ABC}(210) + P_{ABC}(131)P_{ABC}(211) + P_{ABC}(131)P_{ABC}(212)+ \\
	P_{ABC}(131)P_{ABC}(213) + P_{ABC}(131)P_{ABC}(310) + P_{ABC}(131)P_{ABC}(311) + P_{ABC}(131)P_{ABC}(312) + P_{ABC}(131)P_{ABC}(313)+ \\
	P_{ABC}(132)P_{ABC}(201) + P_{ABC}(132)P_{ABC}(210) + P_{ABC}(132)P_{ABC}(211) + P_{ABC}(132)P_{ABC}(212) + P_{ABC}(132)P_{ABC}(213)+ \\
	P_{ABC}(132)P_{ABC}(301) + P_{ABC}(132)P_{ABC}(310) + P_{ABC}(132)P_{ABC}(312) + P_{ABC}(132)P_{ABC}(313) + P_{ABC}(133)P_{ABC}(200)+ \\
	P_{ABC}(133)P_{ABC}(211) + P_{ABC}(133)P_{ABC}(212) + P_{ABC}(133)P_{ABC}(213) + P_{ABC}(133)P_{ABC}(220) + P_{ABC}(133)P_{ABC}(230)+ \\
	P_{ABC}(133)P_{ABC}(300) + P_{ABC}(133)P_{ABC}(311) + P_{ABC}(133)P_{ABC}(312) + P_{ABC}(133)P_{ABC}(313) + P_{ABC}(133)P_{ABC}(320)+ \\
	P_{ABC}(133)P_{ABC}(330) + P_{ABC}(200)P_{ABC}(203) + P_{ABC}(200)P_{ABC}(213) + P_{ABC}(200)P_{ABC}(223) + P_{ABC}(200)P_{ABC}(233)+ \\
	P_{ABC}(200)P_{ABC}(303) + P_{ABC}(200)P_{ABC}(313) + P_{ABC}(200)P_{ABC}(323) + P_{ABC}(200)P_{ABC}(333) + P_{ABC}(203)P_{ABC}(210)+ \\
	P_{ABC}(203)P_{ABC}(220) + P_{ABC}(203)P_{ABC}(230) + P_{ABC}(203)P_{ABC}(300) + P_{ABC}(203)P_{ABC}(310) + P_{ABC}(203)P_{ABC}(320)+ \\
	P_{ABC}(203)P_{ABC}(330) + P_{ABC}(210)P_{ABC}(213) + P_{ABC}(210)P_{ABC}(223) + P_{ABC}(210)P_{ABC}(233) + P_{ABC}(210)P_{ABC}(303)+ \\
	P_{ABC}(210)P_{ABC}(313) + P_{ABC}(210)P_{ABC}(323) + P_{ABC}(210)P_{ABC}(333) + P_{ABC}(213)P_{ABC}(220) + P_{ABC}(213)P_{ABC}(230)+ \\
	P_{ABC}(213)P_{ABC}(300) + P_{ABC}(213)P_{ABC}(310) + P_{ABC}(213)P_{ABC}(320) + P_{ABC}(213)P_{ABC}(330) + P_{ABC}(220)P_{ABC}(223)+ \\
	P_{ABC}(220)P_{ABC}(233) + P_{ABC}(220)P_{ABC}(303) + P_{ABC}(220)P_{ABC}(313) + P_{ABC}(220)P_{ABC}(323) + P_{ABC}(220)P_{ABC}(333)+ \\
	P_{ABC}(223)P_{ABC}(230) + P_{ABC}(223)P_{ABC}(300) + P_{ABC}(223)P_{ABC}(310) + P_{ABC}(223)P_{ABC}(320) + P_{ABC}(223)P_{ABC}(330)+ \\
	P_{ABC}(230)P_{ABC}(233) + P_{ABC}(230)P_{ABC}(303) + P_{ABC}(230)P_{ABC}(313) + P_{ABC}(230)P_{ABC}(323) + P_{ABC}(230)P_{ABC}(333)+ \\
	P_{ABC}(233)P_{ABC}(300) + P_{ABC}(233)P_{ABC}(310) + P_{ABC}(233)P_{ABC}(320) + P_{ABC}(233)P_{ABC}(330) + P_{ABC}(300)P_{ABC}(303)+ \\
	P_{ABC}(300)P_{ABC}(313) + P_{ABC}(300)P_{ABC}(323) + P_{ABC}(300)P_{ABC}(333) + P_{ABC}(303)P_{ABC}(310) + P_{ABC}(303)P_{ABC}(320)+ \\
	P_{ABC}(303)P_{ABC}(330) + P_{ABC}(310)P_{ABC}(313) + P_{ABC}(310)P_{ABC}(323) + P_{ABC}(310)P_{ABC}(333) + P_{ABC}(313)P_{ABC}(320)+ \\
	P_{ABC}(313)P_{ABC}(330) + P_{ABC}(320)P_{ABC}(323) + P_{ABC}(320)P_{ABC}(333) + P_{ABC}(323)P_{ABC}(330) + P_{ABC}(330)P_{ABC}(333)
\end{align*}