### 2016-12-07 figures/causal_structures/triangle_scenario_anim_1.tex

\begin{tikzpicture}[scale=1]
    \begin{scope}[every node/.style=observed, onslide=<2->{fade}]
        \node[onslide=<3>{unfade}] (C) at (-2, 0) {$\p{C}$};
        \node (B) at (2, 0) {$\p{B}$};
        \node[onslide=<2>{unfade}] (A) at (0, {2*sqrt(3)}) {$\p{A}$};
    \end{scope}
    \begin{scope}[every node/.style=latent, onslide=<2->{fade}]
        \node[onslide=<2-3>{unfade}] (X) at (-1, {sqrt(3)}) {$\p{X}$};
        \node[onslide=<2>{unfade}] (Y) at (1, {sqrt(3)}) {$\p{Y}$};
        \node[onslide=<3>{unfade}] (Z) at (0, 0) {$\p{Z}$};
    \end{scope}
    \begin{scope}[every path/.style={draw=cause, thick}, onslide=<2->{fade}]
        \path[postaction={on each segment={mid arrow}}]
        (Y) -- (B)
        (Z) -- (B)
        ;
        \path[postaction={on each segment={mid arrow}}, onslide=<3>{unfade}]
        (X) -- (C)
        (Z) -- (C)
        ;
        \path[postaction={on each segment={mid arrow}}, onslide=<2>{unfade}]
        (X) -- (A)
        (Y) -- (A)
        ;
    \end{scope}
\end{tikzpicture}