### 2017-09-19 old/documents/BrazilSlides/Archive Versions/slides_for_rob.tex

\documentclass[]{beamer}

\usetheme[height=8mm]{Rochester} % Main Theme
\usecolortheme{default} % Color Theme
\usepackage{lmodern} % Optional to remove font size warnings

\begin{document}

\begin{frame}
    \frametitle{Postdoc Opportunities At The Perimeter Institute}
    \begin{center}
    \textbf{Two Postdoctoral Fellowships in Quantum Foundations at the Perimeter Institute} \\
    Project: \textit{Quantum Causal Structures} \\
    \vfill
    \begin{itemize}
        \item How to define quantum causal models
        \item Quantum causal inference
        \item How to provide causal explanations of Bell inequality violations
        \item Exploring the possibilities for indefinite causal structure
    \end{itemize}
    \vfill
    {\scriptsize Application \& Info:}\\

    The Perimeter Website $\to$ Research $\to$ Careers $\to$ Positions $\to$ Quantum Causal Structures Postdoctoral Fellowship
    \textcolor{blue}{\url{https://www.perimeterinstitute.ca/2016/17-quantum-causal-structures-postdoctoral-fellowship}}
    \vfill

    {\scriptsize Funded by the John Templeton Foundation}
    \end{center}

\end{frame}

\end{document}