### 2017-09-19 old/figures/causal_structures/inflated_triangle_scenario_spiral_anim_1.tex

\newcommand{\ift}{2.3}
\begin{tikzpicture}[scale=2]
    \begin{scope}[every node/.style=observed, onslide=<2->{fade}]
        \node[onslide=<3>{unfade}] (C2) at ({-2 + 2*1/\ift}, {2*1/(\ift*sqrt(3))}) {$\p{C}_2$};
        \node (C1) at ({-2 + 3*1/\ift}, {3*1/(\ift*sqrt(3))}) {$\p{C}_1$};
        \node (B2) at ({2 - 2*1/\ift}, {2*1/(\ift*sqrt(3))}) {$\p{B}_2$};
        \node (B1) at ({2 - 3*1/\ift}, {3*1/(\ift*sqrt(3))}) {$\p{B}_1$};
        \node (A2) at (0, {2*sqrt(3) - 2*2/sqrt(3)*(1/\ift)}) {$\p{A}_2$};
        \node[onslide=<2>{unfade}] (A1) at (0, {2*sqrt(3) - 3*2/sqrt(3)*(1/\ift)}) {$\p{A}_1$};
    \end{scope}
    \begin{scope}[every node/.style=latent, onslide=<2->{fade}]
        \node[onslide=<3>{unfade}] (X2) at (-1, {sqrt(3)}) {$\p{X}_2$};
        \node[onslide=<2>{unfade}] (X1) at ({-1 + 1/\ift}, {sqrt(3) - 1/(\ift*sqrt(3))}) {$\p{X}_1$};
        \node (Y2) at (1, {sqrt(3)}) {$\p{Y}_2$};
        \node[onslide=<2>{unfade}] (Y1) at ({1 - 1/\ift}, {sqrt(3) - 1/(\ift*sqrt(3))}) {$\p{Y}_1$};
        \node[onslide=<3>{unfade}] (Z1) at (0, 0.5) {$\p{Z}_1$};
        \node (Z2) at (0, 0) {$\p{Z}_2$};
    \end{scope}
    \begin{scope}[every path/.style={draw=cause, thick}, onslide=<2->{fade}]
        \path[postaction={on each segment={mid arrow}}]
        (Y1) -- (B1) (Y1) -- (B2) (Z1) -- (B1) (Z2) -- (B2)
        ;
        \path[postaction={on each segment={mid arrow}}]
        (X1) -- (A2) (X1) -- (C1) (Z1) -- (C1) (Y2) -- (A2)
        ;
        \path[postaction={on each segment={mid arrow}}, onslide=<3>{unfade}]
        (X2) -- (C2) (Z1) -- (C2)
        ;
        \path[postaction={on each segment={mid arrow}}, onslide=<2>{unfade}]
        (Y1) -- (A1) (X1) -- (A1)
        ;
    \end{scope}
\end{tikzpicture}