### 2017-09-21 CCIAQVTS/figures_source/causal_structures/triangle_structure_injectable_highlight_1.tex
updates for arxiv

\documentclass[preview]{standalone}

\usepackage[showframe]{geometry}
\geometry{
paperwidth=8cm,
paperheight=4cm,
margin=0cm
}
\usepackage{../../document_config}
\input{formatting}
\begin{document}
\begin{tikzpicture}[scale=1]
    \begin{scope}[every node/.style=observed, fade]
        \node         (C) at (-2, 0) {$C$};
        \node         (B) at (2, 0) {$B$};
        \node[unfade] (A) at (0, {2*sqrt(3)}) {$A$};
    \end{scope}
    \begin{scope}[every node/.style=latent, fade]
        \node[unfade] (X) at (-1, {sqrt(3)}) {$X$};
        \node[unfade] (Y) at (1, {sqrt(3)}) {$Y$};
        \node         (Z) at (0, 0) {$Z$};
    \end{scope}
    \begin{scope}[every path/.style={draw=cause, thick}, fade]
        \path[postaction={on each segment={mid arrow}}]
        (Y) -- (B)
        (Z) -- (B)
        ;
        \path[postaction={on each segment={mid arrow}}]
        (X) -- (C)
        (Z) -- (C)
        ;
        \path[postaction={on each segment={mid arrow}}, unfade]
        (X) -- (A)
        (Y) -- (A)
        ;
    \end{scope}
\end{tikzpicture}
\end{document}