### 2018-08-01 documents/Proof Response/proof_response.tex
Made changes to the figures in order to have the figure files match their appearance in the document.

\documentclass{article}
\usepackage{verbatim}
\newenvironment{metaverbatim}{\verbatim}{\endverbatim}

\usepackage[margin=1in]{geometry}

\begin{document}
    \begin{center}
        \textbf{Proof Corrections for AN11411} \\
        \textbf{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure} \\
        Thomas C. Fraser and Elie Wolfe \\
        Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5 \\
        University of Waterloo, Waterloo, Ontario, Canada, N2L 3G1\\
        August 1, 2018
    \end{center}

    \textbf{Specific Questions and Comments to Address for This Paper:}
    \begin{enumerate}
        \item
        \item The authors agree completely with all associated modifications made to the article pursuant to the policy on claims of priority and novelty of contributions.
        \item
        \item The authors approve of all substitutions made regarding the use of slash between words throughout the document.
        \item Thank you for pointing out this one sentence paragraph. The proof matches exactly how the authors would/will correct this mistake.
        \item Noted and approved.
        \item Regarding figures 4, 5, 6 and 7, the incorrect figures were used in the proof. The discrepancy between the figure files and appearance in the submitted article were a result of using the a clipping/trimming option of \verb|\includegraphics| command of the \verb|graphicx| package in order to trim whitespace and superfluous figure titles. The usage of the this option is as follows:
        \begin{verbatim}\includegraphics[trim={<left> <lower> <right> <upper>},clip]{figure_example.pdf}\end{verbatim}
        Attached to this proof correction are ``corrected'' figure replacements for figures 4, 5, 6 and 7 which have the cropping already applied. As such, the figures should now appear correctly in the article and will match the figure files exactly. The figure files are respectively named:
        \begin{verbatim}
        figure_fritz_dist_plotted_bits.pdf
        figure_noisy_yet_non_local_fritz.pdf
        figure_plotted_dist_I_1_max_violation.pdf
        figure_plotted_dist_I_3_max_violation.pdf
        \end{verbatim}
        Regarding figures 4 and 5 only, the colored gradient legends have been removed and replaced with the explicit legends. In the original manuscript, these explicit legends were included adjacent to the figure as an equation. These legends are now a part of the attached figure files, as they should be.
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item
        \item[FQ.]
        \item[Q.]
    \end{enumerate}

\end{document}